CS6213 Project Report

### 2021-05-07 report/paper.tex
Formatting
More

\subsubsection{Extending Specifications}

In our experience, it is fairly easy to specify complex behaviours when you are writing the specification from scratch, but it is a hassle to \emph{extend} an existing spec, \eg, to add history variables to Raft. 
%
TLA specifications are very non-compositional, and even small extensions can require non-local changes.

@@

\subsubsection{Runtime Errors}

On several ocassions, we encountered (trivial) runtime errors after considerable model-checking time (> 15 minutes).
%
We know this is an issue in practice, and not just for us, because the specification of one-at-a-time membership changes for Raft that we adapted had two such issues which, with the default TLC options, would have taken on the order of hours to manifest.
%
This class of problem would be solved by the addition of a type system to TLA.

@@

The Apalache symbolic model checker~\cite{konnov_tla+_2019} does in fact add a type system to TLA.
%
Sadly, this is experimental at this stage, and we encountered several bugs related to type checking and type inference in Apalache.
%
We spent multiple days adding types to the Raft TLA specification and interpreting Apalache's obscure error messages with Igor Konnov's help, but sadly were disappointed by the symbolic model checker's performance.
%
The bad performance matches what the authors report in the paper for model checking, but we had hoped that the SMT solvers Apalache uses are more suited for goal-directed search (for test case generation) than TLC's naive breadth-first search.
%
This turned out not to be the case.

@@

\subsection{Constraining TLC's Search Space}

We found four ways to cut down the state space in TLC:

@@

>>>>

\paragraph{Extending Specifications}
%
In our experience, it is fairly easy to specify complex behaviours when you are writing the specification from scratch, but it is a hassle to \emph{extend} an existing spec, \eg, to add history variables to Raft. 
%
TLA specifications are very non-compositional, and even small extensions can require non-local changes.

@@

\paragraph{Runtime Errors}
%
On several ocassions, we encountered (trivial) runtime errors after considerable model-checking time (> 15 minutes).
%
We know this is an issue in practice, and not just for us, because the specification of one-at-a-time membership changes for Raft that we adapted had two such issues which, with the default TLC options, would have taken on the order of hours to manifest.
%
This class of problem would be solved by the addition of a type system to TLA.

@@

The Apalache symbolic model checker~\cite{konnov_tla+_2019} does in fact add a type system to TLA.
%
Sadly, this is experimental at this stage, and we encountered several bugs related to type checking and type inference in Apalache.
%
We spent multiple days adding types to the Raft TLA specification and interpreting Apalache's obscure error messages with Igor Konnov's help, but sadly were disappointed by the symbolic model checker's performance.

The bad performance matches what the authors report in the paper for model checking, but we had hoped that the SMT solvers Apalache uses are more suited for goal-directed search (for test case generation) than TLC's naive breadth-first search.
%
This turned out not to be the case.

@@

\subsection{Constraining TLC's State Space}

Disappointed with Apalache's performance, we turned back to TLC and started exploring all the available documentation for ways to constrain the state space.
%
Overall, we found four ways to cut down the state space that TLC explores:

@@

We found state constraints and views to be the most powerful techniques, and we used them extensively to guide TLC towards finding interesting traces.

Symmetry sets and action constraints can be implemented in terms of state constraints and views.
%
We used TLC's built-in support for symmetry sets, but found expressing state constraints on history variables to be more ergonomic than writing action constraints.
